# Reproducible Research: Peer Assessment 1

```{r setoptions, message=FALSE, echo=TRUE}
  require(knitr)
  opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data
```{r load}
  # The compressed zip archive
  zip_filename <- "activity.zip"

  # The data file stored inside the compressed zip archive
  data_filename <- "activity.csv"

  # Reading the data
  unzip(zip_filename)
  obs <- read.csv(data_filename)

  # Aggregate by day
  steps_per_day <-   aggregate(steps ~ date, data = obs, sum)
  u_steps_per_day <- aggregate(steps ~ date, data = obs, mean)
  names(u_steps_per_day)[names(u_steps_per_day)=="steps"] <- "u_steps"

  # Aggregate by time and average
  u_steps_per_int <- aggregate(steps ~ interval, data = obs, mean)
   
  # Summarize the data
  summary(obs)
```

Observations: `r dim(obs)[1]`  


## What is mean total number of steps taken per day?
```{r daily_mean}
  mean_steps_per_day <- mean(steps_per_day$steps)
  median_steps_per_day <- median(steps_per_day$steps)
  
  hist(steps_per_day$steps, main="Total Steps per day", 
       xlab="Daily Steps")
  abline(v=mean_steps_per_day, lty = 4, lwd = .1, col = "red", pch = 3)
  abline(v=median_steps_per_day, lty = 2, lwd = .1, col = "blue")
  legend("topright", c(paste("mean =",
                             format(mean_steps_per_day,digits=0,scientific=FALSE)),
                       paste("median =",median_steps_per_day)
                       ), 
         col = c("red", "blue"), lty = c(4, 2),
         text.col=c("red", "blue"), bg = "gray90")
```

## What is the average daily activity pattern?
```{r interval_mean}
  plot(u_steps_per_int$interval, u_steps_per_int$steps, 
       main="Average Steps per Interval", 
       xlab="Five minute Interval Identifier", 
       ylab="Steps", 
       type="l",
       col="blue")

  maxes <- u_steps_per_int[u_steps_per_int$steps == max(u_steps_per_int$steps), ]
  n_maxes <- dim(maxes)[1]

  if (n_maxes == 1) {
    legend <- c(paste("Maximum mean steps:", format(maxes$steps,digits=0)),
                paste("For interval ID:", maxes$interval))
  } else {
    legend <- c(paste("Maximum mean steps:", format(maxes$steps[1],digits=0)),
                paste("For interval IDs:", paste(maxes$interval, collapse=" ")))
  }
  legend("topright", legend, bg = "gray90", text.col="blue")
```


## Imputing missing values
```{r impute}
  # plyr::join is needed for order preservation; otherwise merge() would do
  require(plyr)

  is_missing <- ! complete.cases(obs)
  n_missing <- sum(is_missing)

  # Missing steps get the interval means - assumes all desired days are present w/ NAs
  int_lookup <- u_steps_per_int
  names(int_lookup)[names(int_lookup)=="steps"] <- "u_steps"
  imputers <- join(obs[is_missing,], int_lookup, by="interval")
  imputed_obs <- obs
  imputed_obs$steps[is_missing] <- imputers$u_steps

```


## Are there differences in activity patterns between weekdays and weekends?
  # COmparable dates
  #obs$datePOSIX <- as.POSIXlt(obs$date)
  #obs$datestring <- as.character(obs$date)
